---
title: "3.5 report excercise"
author: "Elias Köppel"
date: "2023-03-05"
output: html_document
---


1) Manually clean the data in the tab "Database S1" and save it as a CSV file that can be read into R.
```{r, warning=FALSE}

#load library
library(tidyverse)

#read in the data, which has been cleaned before manually by deleting data which is not of interest
data1 <- read.csv("C:/Users/elias/OneDrive/Uni/4. Semester/Geo/AGDS I/agds_report_eliaskoeppel/data/data_report_exercise_3.5.csv", header=TRUE, stringsAsFactors=FALSE, sep = ";")

#remove the NA values from the data
data1 <- na.omit(data1) 

```


2) In R, aggregate data per experiment and calculate the log-response ratio (llr) within each experiment.

```{r}

library(knitr)

lrr_per_experiment <- data1 |>
  
  #calculate the log-response ratio for all the measurements
  mutate(lrr = log(increased.CO2/ambient.CO2)) |>
  
  #aggregate data per experiment
  group_by(Experiment) |>
  
  #calculate the mean log-response ration within each experiment
  summarise(lrr = mean(lrr))


#results are being presented as a table
kable(lrr_per_experiment)


```



3) Aggregate data across all experiments for different years since the start of the experiment, distinguishing an early phase (<3 years since start), a mid-phase (3-6 years since start), and a late phase (>6 years since start). Calculate the log-response ratio for each phase


```{r}


library(knitr)

#first step: aggregate data1 to the three categories early-phase, mid-phase and late phase:



  # these values are used to define the different categories 
  values = c(3, 6)


  categories <- data1 |>
  
    #create first category (< 3 years)
    filter(Time..years. < values[1]) |>
    mutate(phase = "early phase") |> 
   bind_rows(
      data1 |>
      
      #create second category (> 3 years and < 6 years)
      filter(Time..years. < values[2] & Time..years. > values[1] ) |>
      mutate(phase = "mid-phase") |>
      bind_rows(
        data1 |>
           
          #create third category (> 6 years)
          filter( Time..years. > values[2] ) |>
          mutate(phase = "late phase")))




#second step: calculate the log-response ratio for each phase: 

lrr_categories <- categories |>
  mutate(lrr = log(increased.CO2/ambient.CO2)) |>
  group_by(phase) |>
  summarise(lrr = mean(lrr))
        


kable(lrr_categories)

```


